A Rough Guide to Pre-processing High-Frequency Animal Tracking Data

### 2020-10-13 05_calibration_data.Rmd
redo residence patch

# get recurse data for a 10m radius
recurse_stats <- getRecursions(data_recurse,
                               radius = 100, timeunits = "mins")

@@

@@

```{r}
fig_calib_residence <- 
  ggplot()+
  geom_point(data = data_recurse,
             aes(x, y, col = res_time),
             shape = 19, alpha = 0.5,
             show.legend = F)+
  scale_colour_distiller(palette = "Greens", direction = 1,
                         trans = "log10",
                         labels = scales::comma,
                         limits = c(5, NA),
                         na.value = "grey")+
  ggspatial::annotation_scale(location = "br")+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "grey99"))+
  coord_sf(crs = 32631)
```

@@

# on known residence points
patch_res_known <- atl_res_patch(data_recurse_subset, 
                                buffer_radius = 5,
                                lim_spat_indep = 100,
                                lim_time_indep = 30,
                                min_fixes = 3)
```

@@

```{r}
# for the known and unkniwn patches
patch_sf_data <- atl_patch_summary(patch_res_known, which_data = "spatial")

@@

@@

```{r}
# patch with residence points and all patches
fig_basic_residence <-
  ggplot()+
  geom_point(data = data_recurse,
            aes(x, y,
                col = res_time > 5),
            show.legend = F,
            size = 1)+
  geom_path(data = patch_summary_data,
            aes(x_median, y_median),
            col = pal[4], lty = 1)+
  geom_point(data = patch_summary_data,
             aes(x_median, y_median),
             size = 10,
             stroke = 2,
             col = pal[4],
             shape = 1,
             show.legend = F)+
  scale_colour_manual(values = c("grey", pal[3]))+
  scale_fill_distiller(palette = "RdPu")+
  ggspatial::annotation_scale(location = "tl")+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "grey99"))+
  coord_sf(crs = 32631,
           xlim = c(NA, 650000))
```

@@

```{r}
# get known patch summary
data_res <- data_unproc[stringi::stri_detect(tID, regex = "(WP)"), ]
# filter for boating
data_res <- data_res[between(UTCtime, '2020-08-24 11:13:00', '2020-08-24 12:36:11'), ]

# get waypoint summary
patch_summary_real <- data_res[, list(nfixes_real = .N,
                                      x_median = round(median(x), digits = -2),
                                      y_median = round(median(y), digits = -2),
                                      duration_real = as.numeric(
                                        max(time) - min(time))), 
                               by = "tID"]

# round median coordinate for inferred patches
patch_summary_inferred <- 
  patch_summary_data[, 
                     c("x_median", "y_median", 
                       "nfixes", "duration")
                     ][, `:=`(x_median = round(x_median, digits = -2),
                              y_median = round(y_median, digits = -2))]

# join with respatch summary
patch_summary_compare <- 
  merge(patch_summary_real,
        patch_summary_inferred, 
        on = c("x_median", "y_median"),
        all.x = TRUE, all.y = TRUE)
```

@@

```{r}
# make figure comparing different methods
# figure_patch_duration <-
  ggplot()+
  geom_point(data = patch_summary_compare,
             aes(duration, 
                 duration_real),
             show.legend = F,
            size = 3,
            stroke = 1,
            shape = 21,
            fill = pal[2],
            col = "white")+
  geom_smooth(data = patch_summary_compare,
             aes(duration, 
                 duration_real),
             se = F,
             size = 0.5,
             col = pal[1],
             alpha = 0.2,
             method = "glm",
             show.legend = F)+
  annotate(geom = "text",
           x = 500, y = 320,
           size = 5,
           label = "R^2 == 0.849",
           parse = T)+
  # scale_x_log10(breaks = c(300, 600, 900),
  #               labels = as.integer(c(300, 600, 900) / 60))+
  # scale_y_log10(breaks = c(300, 600, 900),
  #               labels = as.integer(c(300, 600, 900) / 60))+
  # coord_equal(expand = F)+
  ggthemes::theme_few()+
  labs(x = "inferred duration (min)",
       y = "real duration (min)")
```

@@

```{r}
# wrap together
figure_res_patch <-
  wrap_plots(list(fig_basic_residence, figure_patch_duration))+
    plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")") &
  theme(plot.tag = element_text(face = "bold"))

# save figure
ggsave(figure_res_patch,
       filename = "figures/fig_calib_residence_patch.png",
       height = 130, width = 225, units = "mm")
```

>>>>

# get recurse data for a 10m radius
recurse_stats <- getRecursions(data_recurse,
                               radius = 50, timeunits = "mins")

@@

Read in the data.

```{r}
data_recurse <- fread("data/data_calib_recurse.csv")
```

Subset waypoint data.

```{r}
library(stringi)
data_res <- data_unproc[stri_detect(tID, regex = "(WP)")]
```

@@

```{r}
fig_calib_residence <-
  ggplot()+
  geom_point(data = data_res, 
             aes(x, y),
             size = 5, shape = 2,
             alpha = 0.1)+
  geom_point(data = data_recurse,
             aes(x, y, 
                 col = res_time > 5,
                 size = res_time > 5),
             shape = 19, alpha = 0.4,
             show.legend = F)+
  geom_path(data = data_recurse,
             aes(x, y),
            col = "grey", lwd = 0.2)+
  scale_colour_manual(values = c("grey40", pal[3]))+
  scale_size_manual(values = c(0.5, 2))+
  ggspatial::annotation_scale(location = "br")+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "grey99"))+
  coord_sf(crs = 32631)
```

@@

# on known residence points
patch_res_known <- atl_res_patch(data_recurse_subset, 
                                buffer_radius = 10,
                                lim_spat_indep = 100,
                                lim_time_indep = 5,
                                min_fixes = 3)
```

@@

```{r}
# for the known and unkniwn patches
patch_sf_data <- atl_patch_summary(patch_res_known, which_data = "spatial",
                                   buffer_radius = 20)

@@

### Get waypoint centroids

```{r}
# get centroid
data_res_summary <- data_res[, list(x_median = median(x),
                                    y_median = median(y),
                                    t_median = median(time)),
                             by = "tID"]

# now get times 10 mins before and after
data_res_summary[, `:=`(t_min = t_median - (10 * 60),
                        t_max = t_median + (10 * 60))]

# make a list of positions 10min before and after
wp_data <- mapply(function(l, u, mx, my) {
  tmp_data <- data_unproc[inrange(time, l, u)]
  tmp_data[, distance := sqrt((mx - x)^2 + (my - y)^2)]
  # keep within 50
  tmp_data <- tmp_data[distance <= 50, ]
  # get duration
  return(diff(range(tmp_data$time)))
}, data_res_summary$t_min, data_res_summary$t_max,
   data_res_summary$x_median, data_res_summary$y_median,
SIMPLIFY = FALSE)
```

@@

```{r}
# patch with residence points and all patches
fig_basic_residence <-
  ggplot()+
  geom_sf(data = griend)+
  geom_point(data = data_res_summary, 
             aes(x_median, y_median),
             size = 10, shape = 7,
             # stroke = ,
             col = "grey20",
             alpha = 1)+
    # geom_path(data = data_recurse,
    #          aes(x, y),
    #         col = "grey", lwd = 0.2)+
  # geom_text(data = data_res_summary,
  #           aes(x_median, y_median,
  #               label = tID),
  #           # size = 10, shape = 7,
  #           # stroke = ,
  #           nudge_x = 100,
  #           fontface = "bold",
  #           col = "grey20",
  #           alpha = 1)+
  # geom_text(data = patch_summary_data, 
  #           aes(x_median, y_median,
  #               label = patch),
  #           # size = 10, shape = 7,
  #           # stroke = ,
  #           nudge_x = -50,
  #           fontface = "bold",
  #           col = pal[4],
  #           alpha = 1)+
  geom_point(data = data_recurse,
            aes(x, y,
                col = res_time > 5,
                shape = res_time > 5),
            show.legend = F,
            size = 2)+
  geom_sf(data = patch_sf_data,
          fill = pal[4],
          col = pal[4],
          alpha = 0.3)+
  geom_segment(data = patch_summary_data,
            aes(x = x_end, y = y_end,
                xend = shift(x_start, type = "lag"),
                yend = shift(y_start, type = "lag")),
            col = pal[4], lty = 1,
            lwd = 0.5)+
  # geom_point(data = patch_summary_data,
  #            aes(x_median, y_median),
  #            size = 5,
  #            stroke = 1,
  #            col = pal[4],
  #            shape = 1,
  #            show.legend = F)+
  scale_colour_manual(values = c("grey40", pal[3]))+
  scale_shape_manual(values = c(1, 19))+
  scale_fill_distiller(palette = "RdPu")+
  ggspatial::annotation_scale(location = "tl")+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "grey99"))+
  coord_sf(crs = 32631,
           expand = F,
           ylim = c(5901700, 5903150),
           xlim = c(649300, 650700))
```

@@

```{r}
# get known patch summary
data_res <- data_unproc[stringi::stri_detect(tID, regex = "(WP)"), ]

# get waypoint summary
patch_summary_real <- data_res[, list(nfixes_real = .N,
                                      x_median = round(median(x), digits = -2),
                                      y_median = round(median(y), digits = -2)), 
                               by = "tID"]

# add real duration
patch_summary_real[, duration_real := unlist(wp_data)]
# round median coordinate for inferred patches
patch_summary_inferred <- 
  patch_summary_data[, 
                     c("x_median", "y_median", 
                       "nfixes", "duration", "patch")
                     ][, `:=`(x_median = round(x_median, digits = -2),
                              y_median = round(y_median, digits = -2))]

# join with respatch summary
patch_summary_compare <- 
  merge(patch_summary_real,
        patch_summary_inferred, 
        on = c("x_median", "y_median"),
        all.x = TRUE, all.y = TRUE)

patch_summary_compare[!is.na(tID)]

# drop nas
patch_summary_compare <- na.omit(patch_summary_compare)

# drop patch around WP080
patch_summary_compare <- patch_summary_compare[tID != "WP080", ]
```

12 patches are identified where there are no waypoints, while 2 waypoints are not identified as patches. These waypoints consisted of 6 and 15 (WP098 and WP092) positions respectively, and were lost when the data were aggregated to 30 second intervals.

@@

```{r}
# make figure comparing different methods
figure_patch_duration <-
  ggplot()+
  geom_point(data = patch_summary_compare,
             aes(duration, 
                 duration_real),
             show.legend = F,
            size = 3,
            stroke = 2,
            shape = 1,
            col = pal[2])+
  geom_smooth(data = patch_summary_compare,
             aes(duration, 
                 duration_real),
             se = F,
             col = pal[1],
             lwd = 1,
             method = "lm",
             show.legend = F)+
  annotate(geom = "text",
           x = 600, y = 1020,
           size = 5,
           label = "R^2 == 0.98",
           parse = T)+
  scale_x_continuous(
    breaks = c(400, 600, 800, 1000),
    labels = as.integer(c(400, 600, 800, 1000) / 60)
    )+
  scale_y_continuous(
    breaks = c(400, 600, 800, 1000),
    labels = as.integer(c(400, 600, 800, 1000) / 60)
  )+
  coord_cartesian(expand = T,
  #             # xlim = c(400, 1050),
  #             ylim = c(250, 1050)
  )+
  geom_abline(slope = 1, lty = 2,
              col = "black")+
  ggthemes::theme_few()+
  theme(panel.background = element_rect(fill = "grey99"),
        panel.grid.minor = element_line(colour = "grey"))+
  labs(x = "inferred duration (min)",
       y = "real duration (min)")
```

@@

```{r}
# wrap together
figure_res_patch <-
wrap_plots(list(fig_basic_residence, figure_patch_duration),
           ncol = 1,
           design = "AA\nAA\nAA\nAA\nBB")+
  plot_annotation(tag_levels = "a",
                tag_prefix = "(",
                tag_suffix = ")") &
theme(plot.tag = element_text(face = "bold"))

# save figure
ggsave(figure_res_patch,
       filename = "figures/fig_calib_residence_patch.png",
       height = 225, width = 170, units = "mm")
```
